Sparse Matrix Vector Multiplication with Multiple vectors and Multiple Matrices on the MIC Architecture

### 2013-10-18 ieee_paper.tex
Updated paper. Still stuff to do.  I fixed graphs, and grammar.

@@

The core of the technique is the use of swizzling and permutation
features of the vectors in the Intel MIC architecture. A vector of 512
bits is composed of four lanes (sometimes called channels) of 128 bits, 
which are made of four segments of 32 bits. A swizzling operation allows
the reorder of the elements within each lane, while a permutation allows 
changes the order of entire lanes. 
Figure~\ref{fig:swizzling_permutation}
illustrates these two possibilities. Both permutation and swizzling
support common reordering and broadcast. In the code, the ``permute''
function needs to use the {\tt \_mm512\_permute4f128\_epi32}
instruction which is only defined on integers and therefore in place typecasts
are necessary to convert to a vector to/from the integer type.
\todo{I added a comment stating we preferred swizzling.}
(Note that we prefer swizzling to permutation if possible because we believe 
it is cheaper. Many instructions have swizzling capability, while permuation
is its own instruction.} \todo{reference?}

@@

To perform our experiments and analysis, we use multiple kinds of
matrices which are presented in Figure~\ref{fig:spy_plots}. The
Supercompact matrices (Figure~\ref{fig:supercompact}) have been
generated to only have nonzero elements in the first 32 columns of
the matrix. That way we ensure the simplest form of multiplication,
the amount of elements from the {\tt x} vector data transfered from
the memory will be minimum. The Compact matrices
(Figure~\ref{fig:compact}) are generated to have 32 nonzeros per row
centered around the diagonal and it represents the ideal case for most
application relying on sparse matrix vector multiplication. The Random
matrices (Figure~\ref{fig:compact}) see their nonzero element
randomly (uniformly) distributed in the matrix, they represent the
worst case scenario for cache based architecture where the cache
reutilisation is the lowest from one row to the next.

The other two types of matrices are used in derivative stencils in 2D
and 3D RBF-FD calculations (Figure~\ref{fig:rbf2dnorcm} pictures a
$32$-points stencil of a $23^2$ 2D grid and
Figure~\ref{fig:rbf2dnorcm} shows a $32$-points stencil of a $8^3$
grid). We also apply Cuthill-Mckee reordering to all the matrices This
ordering technique is designed to reduce the distance between the non
zeroes and the diagonal, hopefully allowing a better cache
reutilisation. The reordered version of these two matrices can be seen
in Figure~\ref{fig:rbf2drcm} and in Figure~\ref{fig:rbf3drcm}.

>>>>

\todo{Add few sentence with discretization}

@@

The core of the technique is the use of swizzling and permutation
features of the vectors in the Intel MIC architecture. A vector of 512
bits is composed of four lanes (sometimes called channels) of 128 bits, 
which are made of four segments of 32 bits. A swizzling operation allows
the reorder of the elements within each lane, while a permutation allows 
changes the order of entire lanes. 
Figure~\ref{fig:swizzling_permutation}
illustrates these two possibilities. Both permutation and swizzling
support common reordering and broadcast. In the code, the ``permute''
function needs to use the {\tt \_mm512\_permute4f128\_epi32}
instruction which is only defined on integers and therefore in place typecasts
are necessary to convert to a vector to/from the integer type.
\todo{I added a comment stating we preferred swizzling.}
(Note that we prefer swizzling to permutation if possible because we believe 
it is cheaper. Many instructions have swizzling capability, while permuation
is its own instruction. \todo{reference?}

@@

The different matrices used to perform our experiments and analysis
are presented in Figure~\ref{fig:spy_plots}. The
Supercompact matrices (Figure~\ref{fig:supercompact}) have been
generated to only have nonzero elements in the first 32 columns of
the matrix. That way we ensure the simplest form of multiplication,
the number of elements from the {\tt x} vector data transfered from
the memory is be minimum, they are transferred efficiently and remain 
in cache. The Compact matrices
(Figure~\ref{fig:compact}) are generated to have 32 nonzeros per row
centered around the diagonal and it represents the ideal case for many
applications that rely on sparse matrix vector multiplication. The nonzero
elements of the Random
matrices (Figure~\ref{fig:compact}) see their nonzero elements
randomly (uniformly) distributed in the matrix; they represent the
worst case scenario for a cache based architecture where the cache
reutilisation is the lowest from one row to the next.

The other two types of matrices are used in derivative stencils in 2D
and 3D RBF-FD calculations (Figure~\ref{fig:rbf2dnorcm} pictures a
$32$-points stencil of a $23^2$ 2D grid and
Figure~\ref{fig:rbf2dnorcm} shows a $32$-points stencil of a $8^3$
grid). We also apply a Cuthill-Mckee reordering to all the matrices. This
ordering technique is designed to reduce the distance between the 
nonzeros and the diagonal, hopefully allowing a better cache
reutilisation. The reordered version of these two matrices can be seen
in Figure~\ref{fig:rbf2drcm} and in Figure~\ref{fig:rbf3drcm}.
